---
title: "AMS Blanks"
author: "Brett Longworth"
date: "Jan 23, 2017"
output: html_document
---

Summary of AMS blanks (JME and C1) for both machines from Sep 2014 to present. Raw ratio data are from snics_raw, normalized and blank corrected values are from snics_results. I'm still working on pulling out acetanalides that were used as test samples, so for now I'm rejecting data Fm > 0.05. 

## Data getting and munging

```{r warning=FALSE, message=FALSE}
library(amstools)
library(blanks)
library(odbc)
library(dplyr)
library(ggplot2)
library(knitr)
#library(ggthemr)

#ggthemr("fresh")

from <- as.Date('2014-09-01')

blanks <- getBlankData(from)

#Export data as CSV
write.csv(blanks, "blanks.csv")
```


## Summary

Raw data (Raw1412, Raw1412.sd) are means by target of accepted (unflagged) points. Points above 1e-14 removed. normFm, normFm.sd, and normFm.err are the mean normalized Fm, standard deviation of norm Fm, and mean reported error of normalized blanks. BlkCorrFm is the mean blank Fm after normalization and blank correction. This should distribute around 0.

### Summary table

```{r, echo=FALSE, warning=FALSE}

blanksum <- blanks %>%
  group_by(system, type) %>%
  summarize(
    Raw1412 = mean(c1412x, na.rm = TRUE),
    Raw1412.sd = sd(c1412x, na.rm = TRUE),
    normFm = mean(norm_ratio),
    normFm.sd = sd(norm_ratio),
    normFm.err = mean(merr),
    N = n())
kable(blanksum, digits = c(0,0,2,3,4,5,5,6,0))
```

### Boxplots by system and sample type for all data.

```{r echo=FALSE, warning=FALSE}
ggplot(blanks, aes(x = system, y = c1412x)) + 
  geom_boxplot() + facet_grid(. ~ type) + 
  xlab("System") + ylab("Average Raw 14/12 ratio") +
  ylab(expression(paste("Raw 14/12 ratio ( X", 10^{-16},")"))) +
  theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())

ggplot(blanks, aes(x = system, y = norm_ratio)) + 
  geom_boxplot() + facet_grid(. ~ type) + 
  xlab("System") + ylab("Average Normalized Fm") +
  #ylab(expression(paste("Raw 14/12 ratio ( X", 10^{-16},")"))) +
  theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())
```

### Variability of blanks over time.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(blanks, aes(tp_date_pressed, c1412x, color = system)) +
  scale_y_log10() + geom_smooth() + facet_grid(type ~ ., scale = "free") + 
  geom_point() + theme_bw() + 
  theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())

# ggplot(blanks, aes(tp_date_pressed, he1312, color = system)) +
#   scale_y_log10() + geom_smooth() + facet_grid(type ~ ., scale = "free") + 
#   geom_point() + theme_bw() + 
#   theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())

ggplot(blanks, aes(tp_date_pressed, norm_ratio, color = system)) +
  geom_smooth() + facet_grid(type ~ ., scale = "free") +  theme_bw() + 
  geom_point() + 
  theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())
```

### Flagging

#### By system
```{r echo=FALSE, warning=FALSE}
ggplot(blanks, aes(x = system, y = flagged)) + 
  geom_violin() + facet_grid(. ~ type) + 
  xlab("System") + ylab("Fraction of runs flagged") +
  theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())
```

#### Over time

```{r, echo=FALSE, warning=FALSE, message=FALSE}
qplot(tp_date_pressed, flagged, color = system, data = blanks) +
  geom_smooth() + facet_grid(type ~ ., scale = "free") +  theme_bw() + 
  theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank())
```


